On the uniqueness and stability of dictionaries for sparse representation of noisy signals

### 2017-08-07 manuscripts/PNAS/review/point_by_point_response.txt

RESPONSE: We have included in our revised manuscript a more compelling exposition of the significance of our results and their relevance to the general scientific audience, and thank the editor for the opportunity to address this concern shared by both reviewers. [Summarize high level major points in our cover here.  Also, mention we try to flesh this out in the next comment below.]

-----------------
Since the result is an extension of the noise-free result of [Hillar & Sommer (2015)] to a noisy case, the authors should provide better justification for the significance of contributions. The main result seems to be based on an approach similar to [Hillar & Sommer (2015)], and there is little to no discussion of the computability and tightness of the sufficient condition for stability in the main results. Thus, the submission does not seem to be greatly more valuable than [Hillar & Sommer (2015)] and also not conclusive regarding the central problems of sparse linear coding and dictionary learning. That being said, the paper (with the modifications below) would certainly merit publication in very solid journal like IEEE Transactions on Information Theory, where it would be a median paper, in my opinion.
-----------------

RESPONSE:

We agree with the reviewer that a straightforward extension of the approach and noise-free result of [Hillar & Sommer (2015)] to the noisy case — while still a significant advancement in applied mathematics with actual practical implications — would not necessarily merit publication in PNAS on its own. The reviewer, however, seems to have overlooked the fact that our main result (Thm. 1) generalizes the noise-free result of [Hillar & Sommer (2015)] far beyond a simple extension to the noisy case in ways which provide valuable insight into the dictionary learning model. No explicit comparison was made in the manuscript, but a significant deviation from the approach taken in [Hillar & Sommer (2015)] was necessary to prove these new facts. Moreover, our solution (Cor. 2) to the optimization formulation of the dictionary learning problem (Prob. 2) — that of most interest to optimization theorists and practitioners of their methods — is the first of its kind in both the noise-free and noisy domains. We discuss all of these points in more detail in our response to the reviewer’s later “detailed comments”.

@@

-- I think the biggest potential issue for publication in PNAS is that the significance of the result needs more discussion. Specifically, the authors could address why going from the noise-free case to the noisy case is not incremental. PNAS specifically targets submissions that go beyond what could be published in a more specialized journal. Could the authors, for example, contrast the proof technique between the noisy and noise-free cases? Or, perhaps, the practical implications?
-----------------

RESPONSE:

With all due respect to the reviewer, we believe the most compelling practical implication of a noisy result over a noiseless result is plainly evident: never has there been an experiment without noise and an unstable model is therefore useless in practice. We assume the reviewer must instead be concerned with what additional practical “take-away” message our manuscript contains. In brief, it is this: there exists a procedure by which one can determine if one’s proposed solution to Prob. 1 or Prob. 2 is indeed unique. We have updated the manuscript with an explicit statement of this fact.

The reviewer also raises the concern that, regardless of practical implications, our results may amount to an incremental advance over those of [Hillar & Sommer (2015)], requiring only a straightforward extension of their approach. We disagree with this assessment on both counts: our main result (Thm. 1) goes far beyond a simple extension of that in [Hillar & Sommer (2015)] to the noisy case, and our method of proof includes a significant deviation from their approach.

It is understandable from the presentation of our proof of our main result (Thm. 1) that our approach appears on the surface to be a straightforward extension of that taken in [Hillar & Sommer (2015)]. As observed by our second reviewer, however, the extension to the noisy case did indeed require an original combination of several results in the literature. Aside from these required modifications, our proof of (Thm. 1) differs from the approach taken by [Hillar & Sommer (2015)] in a key way that is tucked into the proof of our Main Lemma — more specifically, by the addition of Lemma 5. [More detail on what goes down in Lemma 5?] We have updated the manuscript to make the reader more keenly aware of this. To be clear, the perspective we take on the problem yields the following powerful new conclusions beyond those of a straightforward extension of the noiseless result of [Hillar & Sommer (2015)] to the noisy case:

@@

2) An extension to the case where the number of dictionary elements is unknown: The results of [Hillar & Sommer (2015)] only apply to the case where the matrix B has the same number of columns as A. We forgo this assumption and show that B must have at least as many columns as A and contains (up to noise) a subset of the columns of A. The size of this subset depends on a simple relation between the number of columns of B and the regularity of the support-set hypergraph.

3) No spark condition requirement for the generating matrix A: One of the mathematically significant contributions made by [Hillar & Sommer (2015)] was to forgo the constraint that the recovery matrix B also satisfy the spark condition, in contrast to all previously known uniqueness results which (either explicitly or implicitly) made this assumption. Our proof is powerful enough to show that, in fact, even the matrix A need not satisfy the spark condition to be identifiable! Rather, it need only be injective on the union of subspaces with supports that form sets in a hypergraph satisfying the SIP. We had omitted this point in our original submission of the manuscript to keep things simple, but have decided to include this interesting fact in our resubmission given that a recently approved NSF grant _____ makes use of this fact in its description of methods. This also required us to redefine the restricted matrix lower bound L to be in terms of a hypergraph, which is an interesting object for further study in its own right.

4) Implications for the computational complexity of the dictionary learning problem: If the size of the support set of reconstructing codes is polynomial in m,k then the pigeonholing argument in the proof of Thm. 1 requires only a polynomial number of samples distributed over a polynomial number of supports; thus N is polynomial in m, k. [Chris? How does this have implications..? ANS: use the P1 and P2 => P=NP] This point was only hinted at in the Discussion section of our original submission, but we have included it in the updated manuscript to make clear the power of our approach over that taken in [Hillar & Sommer (2015)].

@@

RESPONSE: Yes, there is indeed. While all existing uniqueness results require a sufficient number of samples be drawn from *each* of {m \choose k} possible supports, we have shown that in fact it is enough that a sufficient number of samples be drawn from every support in a subset of drastically smaller size, provided this subset forms a regular hypergraph satisfying the singleton intersection property. As was pointed out in our manuscript, a support set of size m satisfying these criteria can always be constructed for any k < m (e.g. take the consecutive intervals of length k in some cyclic order on [m]), and in certain cases such a set can be constructed from as few as 2\sqrt(m) supports (e.g. when m = k^2, take the supports to be the rows and columns of [m] arranged in a square grid).

-----------------
In these terms, it is also unclear to me if a theoretical extension of the uniqueness results to the stability results merit publication in PNAS given the "top 10%" and "sufficient general interest" requirements for publication. It definitely is deserving of publication in a specialty journal.

The work presented is analytical and does not rely on data validation.
-----------------

RESPONSE: Both reviewers have raised this concern and we again thank the editor for the opportunity to address it. It seems both reviewers have not noticed the several ways in which our results generalize the modern results of [Hillar & Sommer (2015)] beyond a straightforward extension to the noisy case, not to mention our solution to the optimization Prob. 2, which is the first of its kind and of most interest to practitioners. We have already discussed these points in detail in our response to the first reviewer’s identical comment.

### 2017-08-07 manuscripts/PNAS/review/point_by_point_response.txt

RESPONSE: We have included in our revised manuscript a more compelling exposition of the significance of our results and their relevance to the general scientific audience, and thank the editor for the opportunity to address this concern shared by both reviewers. [Summarize high level major points in our cover here.  Also, mention we try to flesh this out in the next comment below.]

-----------------	
Since the result is an extension of the noise-free result of [Hillar & Sommer (2015)] to a noisy case, the authors should provide better justification for the significance of contributions. The main result seems to be based on an approach similar to [Hillar & Sommer (2015)], and there is little to no discussion of the computability and tightness of the sufficient condition for stability in the main results. Thus, the submission does not seem to be greatly more valuable than [Hillar & Sommer (2015)] and also not conclusive regarding the central problems of sparse linear coding and dictionary learning. That being said, the paper (with the modifications below) would certainly merit publication in very solid journal like IEEE Transactions on Information Theory, where it would be a median paper, in my opinion.  
-----------------

RESPONSE:

We agree with the reviewer that a straightforward extension of the approach and noise-free result of [Hillar & Sommer (2015)] to the noisy case — while still a significant advancement in applied mathematics with actual practical implications — would not necessarily merit publication in PNAS on its own. The reviewer, however, seems to have overlooked the fact that our main result (Thm. 1) generalizes the noise-free result of [Hillar & Sommer (2015)] far beyond a simple extension to the noisy case in ways which provide valuable insight into the dictionary learning model. A significant deviation from the approach taken in [Hillar & Sommer (2015)] was necessary to prove these new facts, although no explicit comparison was made in the manuscript. Moreover, our solution (Cor. 2) to the optimization formulation of the dictionary learning problem (Prob. 2) — that of most interest to optimization theorists and practitioners of their methods — is the first of its kind in both the noise-free and noisy domains. We discuss all of these points in more detail in our response to the reviewer’s later “detailed comments”.

@@

3) No spark condition requirement for the generating matrix A: One of the mathematically significant contributions made by [Hillar & Sommer (2015)] was to forgo the constraint that the recovery matrix B also satisfy the spark condition, in contrast to all previously known uniqueness results which (either explicitly or implicitly) made this assumption. Our proof is powerful enough to show that, in fact, even the matrix A need not satisfy the spark condition to be identifiable! Rather, it need only be injective on the union of subspaces with supports that form sets in a hypergraph satisfying the SIP. We had omitted this point in our original submission of the manuscript to keep things simple, but have decided to include this interesting fact in our resubmission given that a recently approved NSF grant _____ makes use of this fact in its description of methods. This also required us to redefine the restricted matrix lower bound L to be in terms of a hypergraph, which is an interesting object for further study in its own right.

[4) Implications for the computational complexity of the dictionary learning problem: If the size of the support set of reconstructing codes is polynomial in m,k then the pigeonholing argument in the proof of Thm. 1 requires only a polynomial number of samples distributed over a polynomial number of supports; thus N is polynomial in m, k. [Chris? How does this have implications..?] This point was only hinted at in the Discussion section of our original submission, but we have included it in the updated manuscript to make clear the power of our approach over that taken in [Hillar & Sommer (2015)].

We also wish to reiterate that our solution to Prob. 2, that of most interest to practitioners of dictionary learning methods, is the first of its kind in both the noise-free and noisy domains. We recognize that this was not clearly communicated in the Discussion section of our submitted manuscript, which begins: “In this note, we generalized the approach of (18) to prove the stability of unique solutions to Probs. 1 and 2 […]” We have revised our re-submitted manuscript to make it clear that a solution to Problem 2 has never been supplied before in the literature until now [right?].

@@

The paper's writing and composition is clear (with an exception detailed below), concise, and relevant to the topic. The analytical work appears to be correct to the best of my understanding, and the paper's combined use of several results in the literature to extend an existing dictionary learning uniqueness condition to a new dictionary learning stability condition appears to be original.

Some questions that remain and may be worth discussing deal with the practical impact of these results.

When is it possible to determine (numerically) if enough data has been gathered to obtain an accurate estimate of the underlying sparsity dictionary? 
-----------------

RESPONSE: A statement of this very general nature requires that we calculate the probability that all of the conditions of Thm. 1 are satisfied by random data. We have provided an ‘almost certain’ guarantee of this flavor in Cor 3, wherein the sparse vector supports are constrained to form a hypergraph satisfying the SIP. It would be interesting to calculate furthermore the probability that random supports form a hypergraph satisfying the SIP. We think we have cracked open a nice set of interesting combinatorial problems (with practical implications!) to be solved by the community.

### 2017-08-11 manuscripts/PNAS/review/point_by_point_response.txt

The paper finds conditions under which solutions for the sparse linear coding problem are unique. In sparse linear coding, one is given a matrix Y and the problem is to find a dictionary A and sparse coefficient matrix X such that Y=AX, up to a bounded uncertainty. This paper finds conditions under which A and X are unique up to scalings and permutation. Previous works looked at the noise-free case; this work considers the case of "noisy" measurements and proves a stability condition.

GENERAL COMMENTS

Overall, while I think that the manuscript makes a solid contribution, I believe is better suited to a more specialized journal. Unlike many journals, PNAS gives a criterion for significance: that the paper be within the top 10% of publications and should be of interest to a general scientific audience. Without compelling evidence to the contrary from the authors, it seems to me that the submission does not rise to that level.
-----------------

RESPONSE: We have included in our revised manuscript a more compelling exposition of the significance of our results and their relevance to the general scientific audience, and we thank the editor for the opportunity to address this concern shared by both reviewers.

@@

-----------------
Since the result is an extension of the noise-free result of [Hillar & Sommer (2015)] to a noisy case, the authors should provide better justification for the significance of contributions. The main result seems to be based on an approach similar to [Hillar & Sommer (2015)], and there is little to no discussion of the computability and tightness of the sufficient condition for stability in the main results. Thus, the submission does not seem to be greatly more valuable than [Hillar & Sommer (2015)] and also not conclusive regarding the central problems of sparse linear coding and dictionary learning. That being said, the paper (with the modifications below) would certainly merit publication in very solid journal like IEEE Transactions on Information Theory, where it would be a median paper, in my opinion.
-----------------

RESPONSE:

We agree with the reviewer that a straightforward extension of the approach and noise-free result of [Hillar & Sommer (2015)] to the noisy case — while still a significant advancement in applied mathematics with actual practical implications — would not necessarily merit publication in PNAS on its own. The reviewer, however, seems to have overlooked the fact that our main result (Thm. 1) generalizes the noise-free result of [Hillar & Sommer (2015)] far beyond a simple extension to the noisy case in ways which provide valuable insight into the dictionary learning model. A significant deviation from the approach taken in [Hillar & Sommer (2015)] was necessary to prove these new facts. Moreover, our solution (Cor. 2) to the optimization formulation of the dictionary learning problem (Prob. 2) — that of most interest to optimization theorists and practitioners of their methods — is the first of its kind (to our knowledge) in both the noise-free and noisy domains. We discuss all of these points in more detail in our response to the reviewer’s later “detailed comments”.

We have included in our updated manuscript a discussion of the computability of the sufficient conditions (they are, indeed, computable), as per the reviewer’s request. It is true, however, that we have not proven the tightness of these conditions and the manuscript is therefore not *conclusive* regarding the central problems of sparse linear coding. We believe that these sufficient conditions yield strong enough conclusions to stand on their own. Nonetheless, we have included in our revised manuscript some discussion of where the results can be tightened (e.g. Lem. 4). A sample complexity polynomial in (m,k) can be achieved, however, by limiting the size of the support set of the \tilde x_i, and we have revised the manuscript to discuss this fact (which, incidentally, does *not* follow from the results of [Hillar & Sommer (2015)] in the noiseless case). [And for fixed k […]]

@@

RESPONSE:

With all due respect to the reviewer, we believe the most compelling practical implication of a noisy result over a noiseless result is plainly evident: never has there been an experiment without noise and an unstable model is therefore useless in practice. We assume the reviewer must instead be concerned with what additional practical “take-away” message our manuscript contains. In brief, it is this: there exists a procedure by which one can determine if one’s proposed solution to Prob. 1 or Prob. 2 is indeed unique. We have updated the manuscript with an explicit statement of this fact.

The reviewer also raises the concern that, regardless of practical implications, our results may amount to an incremental advance over those of [Hillar & Sommer (2015)], requiring only a straightforward extension of their approach. We disagree with this assessment on both counts: our main result (Thm. 1) goes far beyond a simple extension of that in [Hillar & Sommer (2015)] to the noisy case, and our method of proof includes a significant deviation from their approach.

It is understandable from the presentation of our proof of our main result (Thm. 1) that our approach appears on the surface to be a straightforward extension of that taken in [Hillar & Sommer (2015)]. As observed by our second reviewer, however, the extension to the noisy case did indeed require an original combination of several results in the literature. Aside from these required modifications, our proof of (Thm. 1) differs from the approach taken by [Hillar & Sommer (2015)] in a key way that is tucked into the proof of our Main Lemma — more specifically, by the addition of Lemma 5. We have updated the manuscript to make the reader more keenly aware of this. To be clear, the perspective we take on the problem yields the following powerful new conclusions beyond those of a straightforward extension of the noiseless result of [Hillar & Sommer (2015)] to the noisy case:

1) An extension to the case where the number of dictionary elements is unknown: The results of [Hillar & Sommer (2015)] only apply to the case where the matrix B has the same number of columns as A. We forgo this assumption and show that B must have at least as many columns as A and contains (up to noise) a subset of the columns of A. The size of this subset depends on a simple relation between the number of columns of B and the regularity of the support-set hypergraph.

2) A significant reduction in sample complexity: To identify the n x m generating dictionary A, [Hillar & Sommer (2015)] require that data be sampled from *every* k-dimensional subspace spanned by the m columns of A (that is, {m \choose k} subspaces in total). We show that the data need only be sampled from m subspaces in total (e.g. those supported by consecutive intervals of length k in some cyclic order on [m]) and in some cases as few as 2\sqrt{m} (e.g. when m = k^2, take the supports to be the rows and columns of [m] arranged in a square grid).

3) Implications for the computational complexity of the dictionary learning problem: If the size of the support set of reconstructing codes is polynomial in m,k then the pigeonholing argument in the proof of Thm. 1 requires only a polynomial number of samples distributed over a polynomial number of supports; thus N is polynomial in m, k. This point was only hinted at in the Discussion section of our original submission, but we have included it in the updated manuscript to make clear the power of our approach over that taken in [Hillar & Sommer (2015)]. [For fixed k…]

4) No spark condition requirement for the generating matrix A: One of the mathematically significant contributions made by [Hillar & Sommer (2015)] was to forgo the constraint that the recovery matrix B also satisfy the spark condition, in contrast to all previously known uniqueness results which (either explicitly or implicitly) made this assumption. Our proof is powerful enough to show that, in fact, even the matrix A need not satisfy the spark condition to be identifiable! Rather, it need only be injective on the union of subspaces with supports that form sets in a hypergraph satisfying the singleton intersection property (SIP). We had omitted this point in our original submission of the manuscript to keep things simple, but have decided to include this interesting fact in our resubmission. This also required us to redefine the restricted matrix lower bound L to be in terms of a hypergraph, which is an interesting object for further study in its own right.

We also wish to reiterate that our solution to Prob. 2, that of most interest to practitioners of dictionary learning methods, is to our knowledge the first of its kind in both the noise-free and noisy domains. We recognize that this was not clearly communicated in the Discussion section of our submitted manuscript, which begins: “In this note, we generalized the approach of (18) to prove the stability of unique solutions to Probs. 1 and 2 […]”

@@

Is this what you meant?
-----------------

RESPONSE: Yes, though actually for every S \in E there are *more than* (k-1)\choose{\bar{m},k} vectors with support in S (not “at least”). Thank you for helping us to clarify our statements. We have updated the manuscript to disambiguate these conditions.

-----------------
-- Related point: Is there a condition on \ell in Theorem 1?
-----------------

RESPONSE: Yes, thanks for this question. The condition on \ell is perhaps best understood as a condition on \bar m not being too large (given \ell) in order to guarantee the recovery of at least one dictionary element. We have updated the manuscript to reflect this, and the condition now reads “provided \bar m < \ell m / (\ell - 1)” where previously it read “provided p = \bar m − \ell( \bar m − m ) is positive”.

@@

-----------------
-- The paper could also compare the sample complexity results on N with those given in the noise-free case. For example, Table I in reference (18) gives a comparison of sample complexity requirements for different conditions in the noise-free case. Where would the results of this paper stand in that table?

-- Otherwise, all the math seemed correct.
-----------------

RESPONSE: We appreciate this question and have added a discussion of this comparison to the manuscript (all sample complexities improve in this work) [DO THIS]. Many thanks for your careful review of the mathematics. We have tried and continue to try to be as clear, concise, and correct as we possibly can to elevate this work into the top echelon of applied mathematical theory papers.

[Do comparison here, state that noiseless is a corollary of our noisy and therefore we don’t discuss the comparison in the paper. They have the table only because they needed randomness to get what we get with SIP. We’re willing to add it if you insist, but we think it detracts from the paper..]

@@

Some questions that remain and may be worth discussing deal with the practical impact of these results.

When is it possible to determine (numerically) if enough data has been gathered to obtain an accurate estimate of the underlying sparsity dictionary?
-----------------

RESPONSE: A statement of this very general nature requires that we calculate the probability that all of the conditions of Thm. 1 are satisfied by random data. We have provided an ‘almost certain’ guarantee of this flavor in Cor 3, wherein the sparse vector supports are constrained to form a hypergraph satisfying the SIP. It would be interesting to calculate furthermore the probability that random supports form a hypergraph satisfying the SIP. We think we have cracked open a nice set of interesting combinatorial problems (with practical implications!) to be solved by the community.

[We could have done this but we believe what sets our results apart from the vast majority of results in this field so far is its deterministic nature from which any probabilistic result can be derived, e.g. the ones we give]

@@

-----------------
Are the conditions and constants/metrics provided in this paper feasible to compute?
-----------------

RESPONSE: We have determined explicit forms for the effectively computable constants in Theorem 1. [But they are not feasible to compute, since they require calculating the RIP of all k-wise subsets of the data. Revive NP-hard argument?]

@@

-----------------
Is there any new intuition to be had regarding the necessary sampling of the union of subspaces from this analysis (with respect to the intuition obtained from existing uniqueness results)?
-----------------

RESPONSE: Yes, there is indeed. While all existing uniqueness results require a sufficient number of samples be drawn from *each* of {m \choose k} possible supports, we have shown that it is enough that a sufficient number of samples be drawn from every support in a subset of drastically smaller size, provided this subset forms a regular hypergraph satisfying the SIP. As was pointed out in our manuscript, a support set of size m satisfying these criteria can always be constructed for any k < m (e.g. take the consecutive intervals of length k in some cyclic order on [m]), and in certain cases such a set can be constructed from as few as 2\sqrt(m) supports (e.g. when m = k^2, take the supports to be the rows and columns of [m] arranged in a square grid).

@@

-----------------
In these terms, it is also unclear to me if a theoretical extension of the uniqueness results to the stability results merit publication in PNAS given the "top 10%" and "sufficient general interest" requirements for publication. It definitely is deserving of publication in a specialty journal.

The work presented is analytical and does not rely on data validation.
-----------------

RESPONSE: Both reviewers have raised this concern and we again thank the editor for the opportunity to address it. It seems both reviewers have not noticed the several ways in which our results generalize the modern results of [Hillar & Sommer (2015)] beyond a straightforward extension to the noisy case, not to mention our solution to the optimization Prob. 2, which as far as we know is the first of its kind and of most interest to practitioners. We have already discussed these points in detail in our response to the first reviewer’s identical comment.

### 2017-08-14 manuscripts/PNAS/chazthm_pnas.tex
added todos in manu

\significancestatement{Many natural signals (visual scenery, speech, EEG, etc.) lacking domain-specific formal models can nonetheless be usefully characterized as linear combinations of few elementary waveforms drawn from a large `dictionary'. We give general conditions guaranteeing when such dictionaries are uniquely and stably determined by data. The result justifies the use of this model as a constraint for blind source separation, and helps explain the observed universality of emergent representations in sparse models of certain natural phenomena.
}

\authordeclaration{The authors declare no conflict of interest.}
\correspondingauthor{\textsuperscript{1}To whom correspondence should be addressed. E-mail: chaz@berkeley.edu, chillar@msri.org}

@@

@@

@@

@@

@@

@@

As mentioned in the previous section, Thm.~\ref{DeterministicUniquenessTheorem} is a particular case of a more general result that forgoes the assumption that $\mathcal{H} \subseteq {[m] \choose k}$ is regular and satisfies the SIP. If instead we require only that the stars $\cap h(i)$ intersect at singletons for all $i \leq q$ (assuming that the nodes of $\mathcal{H}$ are labeled in some order of non-increasing degree), we have that $\bar m \geq k|\mathcal{H}| / \deg(1)$ and, provided $\bar m < k|\mathcal{H}| / (\deg(1) - 1)$, the non-empty submatrix $J$ is of size equal to the largest number $p$ satisfying:
\begin{align}\label{pcond}
\sum_{i=\ell}^{m} \deg(i) > (\bar m + 1 - \ell) (\deg(\ell) - 1) \ \ \text{for all } \ell \leq p \leq q.
\end{align}
Specifically, $J$ contains all nodes of degree exceeding $\deg(p)$ and some subset of those with degreee equal to $\deg(p)$. For the benefit of the reader, we do not prove explicitly this more general result below; it can be discerned from how exactly Lemma \ref{NonEmptyLemma} is incorporated into the proof of Lem.~\ref{MainLemma}.
%As mentioned in the previous section, Thm.~\ref{DeterministicUniquenessTheorem} is a particular case of a more general result, which requires a looser set of constraints on the hypergraph $\mathcal{H}$ and which applies to $n \times \bar m$ matrices $\mathbf{B}$ (and $\bar m$-dimensional codes $\mathbf{\bar x}_i$) with $\bar m \neq m$:

@@

In this note, we generalized the approach of \cite{Hillar15} to prove the stability of unique solutions to Probs.~\ref{InverseProblem} and \ref{OptimizationProblem} while significantly reducing the known sample complexity. Our results justify the application of the sparse linear coding model to blind source separation problems, wherein the goal is to infer the generating dictionary and sparse codes from noisy measurements, \eqref{LinearModel}. We also collect a set of useful mathematical tools and basic facts for future research. The main motivation for this work, however, was to understand how seemingly universal representations emerge from sparse coding models fit to natural data by a variety of methodologies. We elaborate on these applications below after discussing theoretical aspects.

@@

@@

@@

\begin{proof}[Proof (sketch) of Cor.~\ref{ProbabilisticCor}]
First, note that if a set of measure spaces $\{(X_{\ell}, \Sigma_{\ell}, \nu_{\ell})\}_{\ell=1}^p$ has that $\nu_{\ell}$ is absolutely continuous with respect to $\mu$ for all $\ell \in [p]$, where $\mu$ is the standard Borel measure on $\mathbb{R}$, then the product measure $\prod_{\ell=1}^p \nu_{\ell}$ is absolutely continuous with respect to the standard Borel product measure on $\mathbb{R}^p$ (e.g.,  \cite{folland2013real}). By Thm.~\ref{robustPolythm}, there is a polynomial that is nonzero whenever $Y$ has a stable $k$-sparse representation in $\mathbb{R}^m$; in particular, this property (stability) holds with probability one.
\end{proof}

\clearpage

\section{$\ell_1$-norm Extension}

\begin{problem}\label{OptimizationProblemL1}
Find a matrix $\mathbf{B}$ and vectors \mbox{$\mathbf{\bar x}_1, \ldots, \mathbf{\bar x}_N \in \mathbb{R}^m$} that solve:
\begin{align}\label{minsum}
\min \sum_{i = 1}^N \|\mathbf{\bar x}_i\|_1 \ \ 
\text{subject to} \ \ \|\mathbf{z}_i - \mathbf{B}\mathbf{\bar x}_i\|_2 \leq \eta, \ \text{for all $i$}.
\end{align}
\end{problem}

>>>>

\significancestatement{Many natural signals (visual scenery, speech, EEG, etc.) lacking domain-specific formal models can nonetheless be usefully characterized as linear combinations of few elementary waveforms drawn from a large ``dictionary". We give general conditions guaranteeing when such dictionaries are uniquely and stably determined by data. The result justifies the use of this model as a constraint for blind source separation, and helps explain the observed universality of emergent representations in sparse models of certain natural phenomena.
}

% \authordeclaration{The authors declare no conflict of interest.}
\correspondingauthor{\textsuperscript{1}To whom correspondence should be addressed. E-mail: chaz@berkeley.edu, chillar@msri.org}

@@

We comment that quite often theorists in algorithm development solve the related $\ell_1$ problem, which can be shown to be applicable to 
solving $\ell_0$ problems in certain cases.  [*** IMPROVE ON THIS ***]

@@

[*** Make notation $L_{2\mathcal{H}}$ ***]

@@

[*** TODO: remove $J$ from this ***]

@@

[*** Explain Weaker than spark condition example ***]

@@

[*** We remark about algorithms ***]

We remark that there is a strong application to theoretical neuroscience.

\begin{corollary}
Unsupervised neurally plausible sparse feature learning through a wiring bottleneck recovers original sensory representations and neural activities.
\end{corollary}

[*** Make clearer / motivate better ***]

[*** Also discuss uniquness of sparse sensory representations more generally, like Gabors ***]

@@

%%%%%%%%%%%%%%
% need to cut maybe: 
%As mentioned in the previous section, Thm.~\ref{DeterministicUniquenessTheorem} is a particular case of a more general result that forgoes the assumption that $\mathcal{H} \subseteq {[m] \choose k}$ is regular and satisfies the SIP. If instead we require only that the stars $\cap h(i)$ intersect at singletons for all $i \leq q$ (assuming that the nodes of $\mathcal{H}$ are labeled in some order of non-increasing degree), we have that $\bar m \geq k|\mathcal{H}| / \deg(1)$ and, provided $\bar m < k|\mathcal{H}| / (\deg(1) - 1)$, the non-empty submatrix $J$ is of size equal to the largest number $p$ satisfying:
%\begin{align}\label{pcond}
%\sum_{i=\ell}^{m} \deg(i) > (\bar m + 1 - \ell) (\deg(\ell) - 1) \ \ \text{for all } \ell \leq p \leq q.
%\end{align}
%Specifically, $J$ contains all nodes of degree exceeding $\deg(p)$ and some subset of those with degreee equal to $\deg(p)$. For the benefit of the reader, we do not prove explicitly this more general result below; it can be discerned from how exactly Lemma \ref{NonEmptyLemma} is incorporated into the proof of Lem.~\ref{MainLemma}.
%%%%%%%%%%%%%%%%

%As mentioned in the previous section, Thm.~\ref{DeterministicUniquenessTheorem} is a particular case of a more general result, which requires a looser set of constraints on the hypergraph $\mathcal{H}$ and which applies to $n \times \bar m$ matrices $\mathbf{B}$ (and $\bar m$-dimensional codes $\mathbf{\bar x}_i$) with $\bar m \neq m$:

@@

[*** ADD:  Ganguli, Deep Learning, constants, tightness, ***]

In this note, we prove the stability of unique solutions to Probs.~\ref{InverseProblem} and \ref{OptimizationProblem} while significantly reducing the known sample complexity. Our results justify the application of the sparse linear coding model to blind source separation problems, wherein the goal is to infer the generating dictionary and sparse codes from noisy measurements, \eqref{LinearModel}. We also collect a set of useful mathematical tools and basic facts for future research. The main motivation for this work, however, was to understand how seemingly universal representations emerge from sparse coding models fit to natural data by a variety of methodologies. We elaborate on these applications below after discussing theoretical aspects.

@@

Deep learning noise issues: \cite{szegedy2013intriguing, goodfellow2014explaining}.

@@

[*** TODO: make journals abbreviated references ***]

@@

\begin{proof}[Proof (sketch) of Cor.~\ref{ProbabilisticCor}]
First, note that if a set of measure spaces $\{(X_{\ell}, \Sigma_{\ell}, \nu_{\ell})\}_{\ell=1}^p$ has that $\nu_{\ell}$ is absolutely continuous with respect to $\mu$ for all $\ell \in [p]$, where $\mu$ is the standard Borel measure on $\mathbb{R}$, then the product measure $\prod_{\ell=1}^p \nu_{\ell}$ is absolutely continuous with respect to the standard Borel product measure on $\mathbb{R}^p$. % for space:  (e.g.,  \cite{folland2013real}). 
By Thm.~\ref{robustPolythm}, there is a polynomial that is nonzero whenever $Y$ has a stable $k$-sparse representation in $\mathbb{R}^m$; in particular, this property (stability) holds with probability one.
\end{proof}

%\clearpage

%\section{$\ell_1$-norm Extension}
%
%\begin{problem}\label{OptimizationProblemL1}
%Find a matrix $\mathbf{B}$ and vectors \mbox{$\mathbf{\bar x}_1, \ldots, \mathbf{\bar x}_N \in \mathbb{R}^m$} that solve:
%\begin{align}\label{minsum}
%\min \sum_{i = 1}^N \|\mathbf{\bar x}_i\|_1 \ \ 
%\text{subject to} \ \ \|\mathbf{z}_i - \mathbf{B}\mathbf{\bar x}_i\|_2 \leq \eta, \ \text{for all $i$}.
%\end{align}
%\end{problem}
